---
title: "Minerva Analysis"
description: | 
  Statistical analysis of bahavioural data from the Minerva project.
date: "`r Sys.Date()`"
author:
  - first_name: "Sydelle"
    last_name: "de Souza"
    url: https://psydelle.github.io
    affiliation: University of Edinburgh
    affiliation_url: https://www.ed.ac.uk
    orcid_id: 0000-0003-0645-5666

# bibliography: references.bib
output: 
  distill::distill_article:
    code_folding: true
    toc: true
    number_sections: true

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
  warning = FALSE,
  message = FALSE,
  comment = "##",
  R.options = list(width = 60))

# Load tidyverse packages
library(tidyverse)
options(digits = 3)
options(ggplot2.discrete.fill = c("#387ADF", "#FF4500", "#50C4ED"))
palette <- c("#387ADF", "#FF4500", "#50C4ED")
plot_theme <- theme(
            title = element_text(face = "bold"),
            axis.line = element_line(colour = "black", lineend = "round"),
            axis.title = element_text(face = "bold"),
            axis.title.x = element_blank(),
            axis.title.y = element_text(face = "bold"),
            axis.text.y = element_text(face = "bold"),
            axis.text.x = element_text(face = "bold"),
            legend.title = element_blank(),
            legend.text = element_text(face = "bold"),
            legend.position = "bottom",
            strip.text.x = element_text(face = "bold"),
            strip.text.y = element_text(face = "bold"),
            strip.background = element_rect(colour = "black", fill = "white", linewidth = 0.8),
            panel.background = element_rect(colour = "black", fill = "white", linewidth = 0.8),
            panel.grid.major = element_line(colour = "grey", linewidth = 0.5),
            panel.spacing = unit(0, "points"),
            plot.background = element_rect(colour = "white", fill = "white"))

library(distill)
library(lme4)
library(xtable)
library(knitr)
library(gtsummary)
library(psych)
library(kableExtra)
```

```{r load-data-human, include=FALSE}
# Load the data from csv
human <- read_csv("human_clean_for_acl.csv", show_col_types = FALSE)

# set levels for condition 
human$Condition <- factor(human$Condition, levels = c("Compositional", "Collocation", "Idiom"))

```



## Preface
The data that we'll be using for the analysis is the cleaned version of the data. It does not include baselines, the verbs with low accuracy and outliers. Those can be found in the main repo. I'm doing this too keep the size of the data manageable as the minerva simulation data is too large and will take too long to run.

## Behavioral Data
This data has been collected by means of an acceptability judgment task. Participants were asked to judge the acceptability of a verb-noun combination by pressing the y or n key. We are interested in the reaction times (RT) and the accuracy of the responses.

The experiment had three conditions: `r levels(human$Condition)`. We expect RTs to be different across conditions.

<p style="text-align: center;"> **Idioms are fastest, compositional in the middle, collocations are slowest** </p>

 We do no expect accuracy to be different across conditions as these are adult L1 speakers of English and the items are somewhat frequency matched.

### Reaction Times

```{r descriptive-stats-human-rt}

# Group by condition and calculate the mean and standard deviation
correct <- human %>% filter(Accuracy == 1)
RT <- describeBy(correct$RT, group = correct$Condition, mat = TRUE) %>% 
  as.data.frame() %>% 
  rename(Condition = group1, 
         Mean = mean, 
         SD = sd, 
         Median = median, 
         Min = min, 
         Max = max, 
         Skewness = skew, 
         Kurtosis = kurtosis) %>% 
  select(Condition, Mean, SD, Median, Min, Max, Skewness, Kurtosis) 
  rownames(RT) <- NULL

kable(RT, caption = "Descriptive Statistics for Reaction Times by Condition")

```

```{r fig-human-rt-density, layout="l-body-outset"}

human %>% 
  ggplot(aes(x = RT, fill = Condition)) +
  geom_density(alpha = 0.5, linewidth = 0.8) +
  plot_theme + labs(title = "Density Plot of RTs by Condition", x = "Reaction Time (ms)", y = "Density")
```

```{r fig-human-rt, layout="l-body-outset", fig.cap="Boxplot of RTs by Condition"}
# boxplot of RTs by condition
human %>%
  ggplot(aes(x = Condition, y = RT, fill = Condition)) +
  geom_boxplot(alpha = 0.9, linewidth = 0.8) + 
  geom_jitter(width = 0.2, height = 0, alpha = 0.2) + 
    stat_summary(fun.y=mean, geom="point", shape=2, size=2, color="black") +
  plot_theme +
  labs(x = "Condition", y = "Reaction Time (ms)") + 
  theme(legend.position = "none")
```


### Accuracy

```{r descriptive-stats-human-accuracy}
# Group by condition and calculate the mean and standard deviation

Accuracy <- describeBy(human$Accuracy, group = human$Condition, mat = TRUE) %>% 
  as.data.frame() %>% 
  rename(Condition = group1, 
         Mean = mean, 
         SD = sd, 
         Median = median, 
         Min = min, 
         Max = max, 
         Skewness = skew, 
         Kurtosis = kurtosis) %>% 
  select(Condition, Mean, SD, Median, Min, Max, Skewness, Kurtosis) 
  rownames(Accuracy) <- NULL
kable(Accuracy, caption = "Descriptive Statistics for Accuracy by Condition")
```

```{r fig-human-accuracy, layout="l-body-outset"}
# barplot of mean accuracy by condition
human %>% ggplot(aes(x = Condition, y = Accuracy, fill = Condition)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge", color = "black", linewidth = 0.8) +
  geom_errorbar(stat = "summary", fun.data = "mean_cl_normal", position = position_dodge(width = 0.9), width = 0.25) +
  plot_theme +
  labs(title = "Accuracy by Condition", x = "Condition", y = "Accuracy") + theme(legend.position = "none")
```

### Frequency

#### Condition-wise Descriptive Statistics for Frequency

```{r descriptive-stats-human-frequency}

# Group by condition and calculate the mean and standard deviation

Frequency <- describeBy(human$Frequency, group = human$Condition, mat = TRUE) %>% 
  as.data.frame() %>% 
  rename(Condition = group1, 
         Mean = mean, 
         SD = sd, 
         Median = median, 
         Min = min, 
         Max = max, 
         Skewness = skew, 
         Kurtosis = kurtosis) %>% 
  select(Condition, Mean, SD, Median, Min, Max, Skewness, Kurtosis) 
  rownames(Frequency) <- NULL
kable(Frequency, caption = "Descriptive Statistics for Frequency by Condition")
```

```{r fig-human-frequency, layout="l-body-outset"}

# density plot of frequency by condition
human %>% ggplot(aes(x = Frequency, fill = Condition)) +
  geom_density(alpha = 0.5, linewidth = 0.8) +
  plot_theme + labs(title = "Density Plot of Frequency by Condition", x = "Frequency", y = "Density") + facet_wrap(~Condition, scales = "free") + theme(legend.position = "none")
```

#### Correlation between RT and Frequency

```{r frequency, layout="l-body-outset"}

# Scatterplot of RT and Frequency
human %>% ggplot(aes(x = log(Frequency), y = RT, color = Condition)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black",linewidth = 1) +
   plot_theme + theme(legend.position = "none") + theme(axis.text.x = element_text()) +
  labs(title = "Scatterplot of RT and Frequency", x = "log(Frequency)", y = "Reaction Time (ms)") + facet_wrap(~Condition, scales = "free") +
  scale_color_manual(values = palette)

```

## Statistical Analysis

We will fit a linear mixed effects model to the data to account for the repeated measures design. We will include random intercepts for participants and verbs. Predictors will be condition and scale(frequency). Only correct responses will be included in the model. Family will be gamma with identity link.

```{r lmer-human, layout="l-body-outset", echo=TRUE}
# Fit the model

# human_model <- glmer(RT ~ Condition + scale(Frequency) + (1|ID) + (1|Verb), data = human %>% filter(Accuracy == 1), family = Gamma(link = "identity"))

```

### anova

```{r anova-human-high-freq, layout="l-body-outset", echo=TRUE}

hfreq <- human %>% filter(Frequency > 28000)

hfreq %>% group_by(Condition) %>% summarise(Mean = mean(RT), SD = sd(RT), Median = median(RT), Min = min(RT), Max = max(RT)) %>% 
  kable(caption = "Descriptive Statistics for High Frequency Items by Condition")

# anova for high frequency

aov(RT ~ Condition + log(Frequency) + Error(ID/Verb), data = hfreq %>% filter(Accuracy == 1)) %>% 
summary()
  

```